---
title: "one2one benchmark"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(microbenchmark)
library(edwards)
```

Benchmarking for different implementations of the one2one() function. 

I started with [this thread](https://stackoverflow.com/questions/52399474/check-if-variables-are-in-a-one-to-one-mapping). There were several alternatives suggested and I've benchmarked those here, together with a dplyr implementation of my own.

```{r data, echo = FALSE}
# Create some test sets: dt is one-to-one, df is not 
n = 1E6 #num of rows
x <- sample(1:26, n, replace = T)
y <- letters[x]
y2 <- letters[sample(1:26, n, replace = T)]
dt <- tibble(x, y)
df <- tibble(x, y = y2)
```

The first function uses `table()` then checks that each row and column only has one non-zero entry. In the thread the function just checks columns, but rows must be checked too.

```{r func1}
one2one1 <- function(df) {
  tt <- table(df)
  test1 <- isTRUE(all.equal(colSums(tt), apply(tt, 2,  max)))
  test2 <- isTRUE(all.equal(rowSums(tt), apply(tt, 1,  max)))
  test1 && test2
}
```

The next function uses `duplicated()`.

```{r func2}
one2one2 <- function(df) {
  all(duplicated(df)|duplicated(df, fromLast = TRUE))
}
```

The next two are my dplyr-based attempts. The first compares the `count()` with both columns grouped and the count with single groupings. The second checks the grouped counts for repeated occurences of entries from either column.

```{r func3}
one2one3 <- function(df) {
  count_all <- group_by_all(df) %>% 
    tally() %>% 
    ungroup()
  count_1 <- group_by_at(df, 1) %>% tally()
  count_2 <- group_by_at(df, 2) %>% tally()
  test1 <- identical(select(count_all, c(1, 3)), count_1)
  test2 <- identical(select(count_all, c(2, 3)), count_2)
  test1 && test2
}

one2one4 <- function(df, ...) {
  counts_all <- select(df, ...)%>%
    group_by_all() %>%
    tally() %>%
    ungroup()
  for (col in 1 : (ncol(counts_all) - 1)){
    max_count <- group_by_at(counts_all, {{col}}) %>%
      tally() %>%
      pull(n) %>%
      max()
    if (max_count > 1){
      return(FALSE)
    }
  }
  TRUE
}
```

All functions give correct output:
```{r checks}
all(one2one1(dt), one2one2(dt), one2one3(dt), one2one4(dt))
any(one2one1(df), one2one2(df), one2one3(df), one2one4(dt))
```

Here are the benchmarks for both true and false tests:

```{r bench}
microbenchmark(f1 = one2one1(dt),
               f2 = one2one2(dt),
               f3 = one2one3(dt),
               f4 = one2one4(dt, x, y),
               times = 10)

microbenchmark(f1 = one2one1(df),
               f2 = one2one2(df),
               f3 = one2one3(df),
               f4 = one2one4(df, x, y),
               times = 10)
```

Here's a look at times for some components of the functions, first the false dataset, then the true dataset.
```{r bench2}
microbenchmark(f1 = table(df),
               f2 = tally(group_by_all(df)),
               f3 = tally(group_by_at(df, 1)),
               times = 10)

microbenchmark(f1 = table(dt),
               f2 = tally(group_by_all(dt)),
               f3 = tally(group_by_at(dt, 1)),
               times = 10)
```
